---
title: "Final Project Submission"
author: "Wesley Lim"
date: "2022-11-20"
output: html_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(reticulate)
library(tidyverse)
library(tidytuesdayR)
```

This dataset deals with untidy, raw data for recording global plastic pollution. [https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-01-26/readme.md]
This data set contains data about global plastic pollution in the years 2019-2020. The data can be categorized by plastic type, country of cleanup, and source of plastic. While the data isn't necessarily an accurate representative of plastic pollution, it can give a glimpse into the situation given increasing environmental concerns.

In this project, I aimed to use R to tidy up the data and present statistical tests and graphical representations. At the same time, I attempted to learn python and use its PANDAS package to perform the same tasks if availiable.

Here, I loaded the dataset and got a sense of what I am working with. Verrry untidy data at first glance.

```{r}
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv')
glimpse(plastics)
```

Here, I created different objects to group the data set by, creating different perspectives.
```{r}
plastics_T_2019 <- plastics %>%
  select(country,grand_total,year) %>%
  filter(year == 2019)


```



```{python}
### Python
print("Hello World")

```
